\\Large The heat is on: Rising temperatures alter how and where mammals move

### 2025-09-09 writing/appendix-b-gams.Rmd
changed scientific names to common names

---
title: "\\Large The heat is on: Rising temperatures alter when, how, and where large terrestrial mammals move"
subtitle: "\\Large Appendix B: Modeling"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [braes, biol]
  - name: Chris H. Fleming
    institute: [ucf, smith]
  - name: Siobhan Darlington
    institute: [braes, biol]
  - name: Adam T. Ford
    institute: [braes, biol]
  - name: TJ Gooliaff
    institute: [forest]
  - name: Karen E. Hodges
    institute: [braes, biol]
  - name: Kirk Safford
    institute: bcp
  - name: Robert Serrouya
    institute: [braes, biol, biodiv]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - forest: British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development, Penticton, BC, Canada.
  - bcp: British Columbia Ministry of Environment and Parks, Penticton, BC, Canada.
  - biodiv: Wildlife Science Centre, Biodiversity Pathways, University of British Columbia Okanagan, Revelstoke, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'bc-mammals-temperature.bib'
csl: 'global-change-biology.csl'
fontsize: 12pt
indent: true
header-includes:
    - \renewcommand{\figurename}{Figure B\!\!} # for "Fig. Bx.
    - \renewcommand{\tablename}{Table B\!\!} # for "Table. Bx.
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for Delta symbol in table
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: true
---

@@

```{r pred-vs-pred-hgams, fig.cap="Hexplot of the fitted values from the HGAMs with and without including temperature for the probability of movement (\\textbf{A}) and speed when moving (\\textbf{B})."}
knitr::include_graphics('../figures/hgam-with-without-temp-prediction-agreement.png')
```

@@

```{r op, fig.cap="Relationsips between the observed and predicted values for probability of movement (\\textbf{A}) and speed given that an animal was moving (\\textbf{B}). The color in panel \\textbf{A} indicates the number of points used to calculate the estimate (on the $\\log_{2}$ scale), while the grey lines in both panels indicate perfect prediction (i.e., the 1:1 line)."}
knitr::include_graphics('../figures/p-and-speed-observed-vs-predicted.png')
```

@@

knitr::kable(temp_comp, digits = 1, format = 'markdown', caption = "Changes in deviance expained ($\\Delta$DE) and Akaike Information Criterion ($\\Delta$AIC) from including temperature in the models estimating probability of movement ($P(M)$) and speed when moving (Speed) as well as in species-level Hierarchical Resource Selection Functions (HRSF). Unlike conventional marginal AIC [@akaike_new_1974], The AIC values provided by the `mgcv` package for `R` are calculated using the models' estimated effective degrees of freedom and the Bayesian variance-covariance matrices, which avoids issues related to random effects or Restricted Maximum Likelihood (REML).")
```

>>>>

---
title: "\\Large The heat is on: Rising temperatures alter how and where mammals move"
subtitle: "\\Large Appendix B: Modeling"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [braes, biol]
  - name: Chris H. Fleming
    institute: [ucf, smith]
  - name: Siobhan Darlington
    institute: [braes, biol]
  - name: Adam T. Ford
    institute: [braes, biol]
  - name: TJ Gooliaff
    institute: [forest]
  - name: Karen E. Hodges
    institute: [braes, biol]
  - name: Kirk Safford
    institute: bcp
  - name: Robert Serrouya
    institute: [braes, biol, biodiv]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - forest: British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development, Penticton, BC, Canada.
  - bcp: British Columbia Ministry of Environment and Parks, Penticton, BC, Canada.
  - biodiv: Wildlife Science Centre, Biodiversity Pathways, University of British Columbia Okanagan, Revelstoke, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'bc-mammals-temperature.bib'
csl: 'global-change-biology.csl'
fontsize: 12pt
indent: true
header-includes:
    - \renewcommand{\figurename}{Figure B\!\!} # for "Fig. Bx.
    - \renewcommand{\tablename}{Table B\!\!} # for "Table. Bx.
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for Delta symbol in table
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: true
---

@@

```{r pred-vs-pred-hgams, fig.cap="Hexplot of the fitted values from the HGAMs with and without including temperature for the probability of moving (\\textbf{A}) and speed when moving (\\textbf{B})."}
knitr::include_graphics('../figures/hgam-with-without-temp-prediction-agreement.png')
```

@@

```{r op, fig.cap="Relationsips between the observed and predicted values for probability of moving (\\textbf{A}) and speed given that an animal was moving (\\textbf{B}). The color in panel \\textbf{A} indicates the number of points used to calculate the estimate (on the $\\log_{2}$ scale), while the grey lines in both panels indicate perfect prediction (i.e., the 1:1 line)."}
knitr::include_graphics('../figures/p-and-speed-observed-vs-predicted.png')
```

@@

knitr::kable(temp_comp, digits = 1, format = 'markdown', caption = "Changes in deviance expained ($\\Delta$DE) and Akaike Information Criterion ($\\Delta$AIC) from including temperature in the models estimating probability of moving ($P(M = 1)$) and speed when moving (Speed) as well as in species-level Hierarchical Resource Selection Functions (HRSF). Unlike conventional marginal AIC [@akaike_new_1974], The AIC values provided by the `mgcv` package for `R` are calculated using the models' estimated effective degrees of freedom and the Bayesian variance-covariance matrices, which avoids issues related to random effects or Restricted Maximum Likelihood (REML).")
```
